Spike-and-Slab Generalized Additive Models and Scalable Algorithms for High-Dimensional Data

### 2022-02-18 Manuscript/R&R/03-reviewer_4.Rmd
Add citations (#25)

_Comments to the authors_

@@

_The authors introduce the Bayesian hierarchical generalized additive model (BHAM) for high-dimensional data analysis with generalized additive models (GAMs). The BHAM framework allows for variable selection and estimation in GAMs. However, in contrast to previous works on Bayesian GAMs, bilevel selection is possible under BHAM, and BHAM allows practitioners to more easily distinguish linear effects from nonlinear effects. Two deterministic algorithms, EM-Coordinate Descent (EM-CD) and EM-Iterative Weighted Least Squares (EM-IWLS), are introduced to fit BHAM. The EM-IWLS algorithm allows for inference, which is its main advantage over EM-CD. The algorithms appear to be faster than GAMs based on group regularization since fitting a model with a lasso penalty is typically faster than fitting one with a group lasso penalty._

_The paper is coherent and introduces a potentially promising methodology that addresses some of the shortcomings of existing Bayesian GAM approaches. However, in order to be suitable for publication in Statistics in Medicine, I have a few concerns that I think the authors should address. Namely, the authors need to do a better job illustrating the benefits of the BHAM approach over existing methods (such as the sparse Bayesian GAM method of [1, 2] or the spikeslabGAM of [3]). I detail my main concerns below._

## _2. Main Comments and Questions_
\em 1. The incorporation of the smoothing penalty matrix Sj in BHAM allows for separation of the linear space of a smoothing function from the nonlinear space. This is one of the touted benefits of BHAM - this separation allows for bilevel selection rather than "all-in-all-out" selection, so that the practitioner can distinguish linear effects from nonlinear ones. However, one concern is that the method might select the nonlinear effects but fail to select the linear component. The authors briefly allude to this in the Conclusion (p. 13 of the manuscript). The authors say that they "[include] the linear component in the model when non-linear component is selected."

@@

For example, in bilevel selection with the sparse group lasso [4], Simon et al. (2012)
state in their algorithm that they first check whether or not the group is selected,
and only if the group is nonzero do they then proceed with regularizing the individual
coordinates in the nonzero groups. Is a similar check being done for BHAM? This
needs to be explained further and made explicit. If necessary, the authors of the
present manuscript could include an additional subsection in Section 2 that explains
how to enforce strong hierarchy (i.e. linear effect is always selected if the corresponding
nonlinear effect for the jth covariate is selected).\em

@@

\em 3. The authors claim that BHAM is computationally beneficial because it does not rely
on MCMC. The authors illustrate the computational benefit of the double exponential
prior over the group spike-and-slab prior of SBGAM [1, 2], but in my view, they
should also compare it to the method spikeslabGAM of [3]. There is an R package
spikeslabGAM to implement this method. This method uses a Gaussian spike-and-
slab prior on the basis coefficients and is implemented using MCMC. In order to assess
the benefit of using double exponential priors in BHAM vs. Gaussian spike-and-slab
priors, as well as illustrating the computational efficiency of the EM-CD and EM-
IWLS algorithms compared to MCMC, the authors should include comparisons to
spikeslabGAM.\em

>>>>

\em Comments to the authors\em

@@

\em The authors introduce the Bayesian hierarchical generalized additive model (BHAM) for high-dimensional data analysis with generalized additive models (GAMs). The BHAM framework allows for variable selection and estimation in GAMs. However, in contrast to previous works on Bayesian GAMs, bilevel selection is possible under BHAM, and BHAM allows practitioners to more easily distinguish linear effects from nonlinear effects. Two deterministic algorithms, EM-Coordinate Descent (EM-CD) and EM-Iterative Weighted Least Squares (EM-IWLS), are introduced to fit BHAM. The EM-IWLS algorithm allows for inference, which is its main advantage over EM-CD. The algorithms appear to be faster than GAMs based on group regularization since fitting a model with a lasso penalty is typically faster than fitting one with a group lasso penalty.

The paper is coherent and introduces a potentially promising methodology that addresses some of the shortcomings of existing Bayesian GAM approaches. However, in order to be suitable for publication in Statistics in Medicine, I have a few concerns that I think the authors should address. Namely, the authors need to do a better job illustrating the benefits of the BHAM approach over existing methods (such as the sparse Bayesian GAM method of [@Bai2020; @Bai2021] or the spikeslabGAM of [@Scheipl2012]). I detail my main concerns below.\em

## _2. Main Comments and Questions_

\em 1. The incorporation of the smoothing penalty matrix $S_j$ in BHAM allows for separation of the linear space of a smoothing function from the nonlinear space. This is one of the touted benefits of BHAM - this separation allows for bilevel selection rather than "all-in-all-out" selection, so that the practitioner can distinguish linear effects from nonlinear ones. However, one concern is that the method might select the nonlinear effects but fail to select the linear component. The authors briefly allude to this in the Conclusion (p. 13 of the manuscript). The authors say that they "[include] the linear component in the model when non-linear component is selected."

@@

For example, in bilevel selection with the sparse group lasso [@simon2013], Simon et al. (2012)
state in their algorithm that they first check whether or not the group is selected,
and only if the group is nonzero do they then proceed with regularizing the individual
coordinates in the nonzero groups. Is a similar check being done for BHAM? This
needs to be explained further and made explicit. If necessary, the authors of the
present manuscript could include an additional subsection in Section 2 that explains
how to enforce strong hierarchy (i.e. linear effect is always selected if the corresponding
nonlinear effect for the jth covariate is selected).\em

@@

_3. The authors claim that BHAM is computationally beneficial because it does not rely
on MCMC. The authors illustrate the computational benefit of the double exponential
prior over the group spike-and-slab prior of [@Bai2020; @Bai2021], but in my view, they
should also compare it to the method spikeslabGAM of @Scheipl2012. There is an R package
spikeslabGAM to implement this method. This method uses a Gaussian spike-and-
slab prior on the basis coefficients and is implemented using MCMC. In order to assess
the benefit of using double exponential priors in BHAM vs. Gaussian spike-and-slab
priors, as well as illustrating the computational efficiency of the EM-CD and EM-
IWLS algorithms compared to MCMC, the authors should include comparisons to
spikeslabGAM._
